Automated Residual Plot Assessment with the R Package autovi and Shiny App autovi.web

### 2024-11-23 autovi_paper.qmd
finish revising paper

The modules for VSS prediction and $p$-value computation are predefined and cannot be overridden, although users can interact with them directly through function arguments. Similarly, the image reading and resizing module is fixed but will adapt to different Keras models by checking their input shapes. The remaining seven modules are designed to be overridable, enabling users to tailor the infrastructure to their specific needs. These modules are discussed in detail on the software's website. <!-- in the following sections.-->

@@

The `autovi.web` package extends the functionality of `autovi` by offering a user-friendly web interface for automated residual plot assessment. This eliminates the common challenges associated with software installation, so users can avoid managing Python environments or handling version requirements for R libraries. The platform is cross-platform and accessible on various devices and operating systems, making it suitable even for users without R programming experience. Additionally, updates are managed centrally, ensuring that users always have access to the latest features.
<!--The `autovi.web` interface is available at [autoviweb.netlify.app](autoviweb.netlify.app).--> This section discusses the implementation based on `autovi.web` version 0.1.0.

@@

The package `autovi.web` is built using the `shiny` [@shiny] and `shinydashboard` [@shinydashboard] R packages. Hosted on the [shinyapps.io](https://www.shinyapps.io) domain, the application is accessible through any modern web browser. The R packages `htmltools` [@htmltools] and `shinycssloaders` [@shinycssloaders] are used to render markdown documentation in shiny application, and for loading animations for shiny widgets, respectively.

@@

This paper presents new regression diagnostics software, the R package `autovi` and its accompanying web interface package, `autovi.web`. It addresses a critical gap in the current landscape of statistical software. While regression tools are widely available, effective and efficient diagnostic methods have lagged behind, particularly in the field of residual plot interpretation.

@@

The the current version of `autovi` can be installed from CRAN, and source code for both packages are available at [https://github.com/TengMCing/autovi](https://github.com/TengMCing/autovi). The web interface is available from [autoviweb.netlify.app](autoviweb.netlify.app).

>>>>

The modules for VSS prediction and $p$-value computation are predefined and cannot be overridden, although users can interact with them directly through function arguments. Similarly, the image reading and resizing module is fixed but will adapt to different Keras models by checking their input shapes. The remaining seven modules are designed to be overridable, enabling users to tailor the infrastructure to their specific needs. These modules are discussed in detail on the software's website [https://tengmcing.github.io/autovi/](https://tengmcing.github.io/autovi/). <!-- in the following sections.-->

@@

The `autovi.web` shiny application extends the functionality of `autovi` by offering a user-friendly web interface for automated residual plot assessment. This eliminates the common challenges associated with software installation, so users can avoid managing Python environments or handling version requirements for R libraries. The platform is cross-platform and accessible on various devices and operating systems, making it suitable even for users without R programming experience. Additionally, updates are managed centrally, ensuring that users always have access to the latest features.
<!--The `autovi.web` interface is available at [autoviweb.netlify.app](autoviweb.netlify.app).--> This section discusses the implementation based on `autovi.web` version 0.1.0.

@@

The interface `autovi.web` is built using the `shiny` [@shiny] and `shinydashboard` [@shinydashboard] R packages. Hosted on the [shinyapps.io](https://www.shinyapps.io) domain, the application is accessible through any modern web browser. The R packages `htmltools` [@htmltools] and `shinycssloaders` [@shinycssloaders] are used to render markdown documentation in shiny application, and for loading animations for shiny widgets, respectively.

@@

This paper presents new regression diagnostics software, the R package `autovi` and its accompanying web interface, `autovi.web`. It addresses a critical gap in the current landscape of statistical software. While regression tools are widely available, effective and efficient diagnostic methods have lagged behind, particularly in the field of residual plot interpretation.

@@

The current version of `autovi` can be installed from CRAN, and source code for both packages are available at [https://github.com/TengMCing/autovi](https://github.com/TengMCing/autovi) and [https://github.com/TengMCing/autovi_web](https://github.com/TengMCing/autovi_web) respectively. The web interface is available from [autoviweb.netlify.app](autoviweb.netlify.app).

### 2024-11-23 autovi_paper.tex
finish revising paper

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

@@

\usepackage[]{natbib}
\bibliographystyle{anzsj}
\usepackage{bookmark}

@@

However relying solely on subjective assessments of these plots can lead
to issues, such as over-interpreting random patterns as model
violations. \citet{li2024plot} demonstrated that visual methods using
the lineup protocol \citep{buja2009statistical} for assessing residuals
are more useful, and also perform more practically than conventional
tests due to their reduced sensitivity to minor departures. Packages
such as \texttt{nullabor} \citep{nullabor}, \texttt{HLMdiag}
\citep{loy2014hlmdiag}, and \texttt{regressinator}
\citep{regressinator}, enable users to compare observed residual plots
with samples from null distributions, helping to quantify the
significance of any detected patterns. However, the

However, as discussed in \citet{li2024automated}, the lineup protocol
has significant limitations in large-scale applications due dependence
on human labor. Thus a computer vision model was developed with an
associated statistical testing procedure to automate the assessment of
residual plots. This model takes a residual plot and a vector of
auxiliary variables (such as the number of observations) as inputs and
outputs the predicted visual signal strength (VSS). This strength
estimates the distance between the residual distribution of the fitted
regression model and the reference distribution assumed under correct
model specification.

@@

Figure~\ref{fig-three-examples} shows three sets of plots of residuals
against fitted values. The simulated example in (a) might be interpreted
as a heteroscedastic pattern, however the automated reading would
predict this to have a visual signal strength (VSS) of 1.53, with a
corresponding \(p\)-value of 0.25. This means it is would be interpreted
as a good residual plot, that there is nothing in the data to indicate a
violation of model assumptions. Skewness in the predictor variables is
generating the apparent heteroscedasticity, where the smaller variance
in residuals at larger fitted values is due to smaller sample size only.
The Breusch-Pagan test \citep{breusch1979simple} for heteroscedasticity
would also not reject this as good residual plot.

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-three-examples-1.pdf}

@@

\includegraphics[width=0.6\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-plot-pair-1.pdf}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-summary-plot-1.pdf}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-diag-1.png}

@@

The modules for VSS prediction and \(p\)-value computation are
predefined and cannot be overridden, although users can interact with
them directly through function arguments. Similarly, the image reading
and resizing module is fixed but will adapt to different Keras models by
checking their input shapes. The remaining seven modules are designed to
be overridable, enabling users to tailor the infrastructure to their
specific needs. These modules are discussed in detail on the software's
website.

@@

The \texttt{autovi.web} package extends the functionality of
\texttt{autovi} by offering a user-friendly web interface for automated
residual plot assessment. This eliminates the common challenges
associated with software installation, so users can avoid managing
Python environments or handling version requirements for R libraries.
The platform is cross-platform and accessible on various devices and
operating systems, making it suitable even for users without R
programming experience. Additionally, updates are managed centrally,
ensuring that users always have access to the latest features. This
section discusses the implementation based on \texttt{autovi.web}
version 0.1.0.

@@

The package \texttt{autovi.web} is built using the \texttt{shiny}
\citep{shiny} and \texttt{shinydashboard} \citep{shinydashboard} R
packages. Hosted on the \href{https://www.shinyapps.io}{shinyapps.io}
domain, the application is accessible through any modern web browser.
The R packages \texttt{htmltools} \citep{htmltools} and
\texttt{shinycssloaders} \citep{shinycssloaders} are used to render
markdown documentation in shiny application, and for loading animations
for shiny widgets, respectively.

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-web-setup-1.png}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-web-lineup-1.png}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{autovi_paper_files/figure-pdf/fig-autovi-web-distributions-1.png}

@@

This paper presents new regression diagnostics software, the R package
\texttt{autovi} and its accompanying web interface package,
\texttt{autovi.web}. It addresses a critical gap in the current
landscape of statistical software. While regression tools are widely
available, effective and efficient diagnostic methods have lagged
behind, particularly in the field of residual plot interpretation.

@@

The the current version of \texttt{autovi} can be installed from CRAN,
and source code for both packages are available at
\url{https://github.com/TengMCing/autovi}. The web interface is
available from \url{autoviweb.netlify.app}.

>>>>

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

@@

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{anzsj}
\usepackage{bookmark}

@@

However relying solely on subjective assessments of these plots can lead
to issues, such as over-interpreting random patterns as model
violations. \citet{li2024plot} demonstrated that visual methods using
the lineup protocol \citep{buja2009statistical} for assessing residuals
are more useful, and also perform more practically than conventional
tests due to their reduced sensitivity to minor departures. Packages
such as \texttt{nullabor} \citep{nullabor}, \texttt{HLMdiag}
\citep{loy2014hlmdiag}, and \texttt{regressinator}
\citep{regressinator}, enable users to compare observed residual plots
with samples from null distributions, helping to quantify the
significance of any detected patterns.

However, as discussed in \citet{li2024automated}, the lineup protocol
has significant limitations in large-scale applications due dependence
on human labor. Thus, a computer vision model was developed with an
associated statistical testing procedure to automate the assessment of
residual plots. This model takes a residual plot and a vector of
auxiliary variables (such as the number of observations) as inputs and
outputs the predicted visual signal strength (VSS). This strength
estimates the distance between the residual distribution of the fitted
regression model and the reference distribution assumed under correct
model specification.

@@

Figure~\ref{fig-three-examples} shows three sets of plots of residuals
against fitted values. The simulated example in (a) might be interpreted
as a heteroscedastic pattern, however the automated reading would
predict this to have a visual signal strength (VSS) of 1.53, with a
corresponding \(p\)-value of 0.25. This means it would be interpreted as
a good residual plot, that there is nothing in the data to indicate a
violation of model assumptions. Skewness in the predictor variables is
generating the apparent heteroscedasticity, where the smaller variance
in residuals at larger fitted values is due to smaller sample size only.
The Breusch-Pagan test \citep{breusch1979simple} for heteroscedasticity
would also not reject this as good residual plot.

@@

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-three-examples-1.pdf}

@@

\includegraphics[width=0.6\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-plot-pair-1.pdf}

@@

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-summary-plot-1.pdf}

@@

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-autovi-diag-1.png}

@@

The modules for VSS prediction and \(p\)-value computation are
predefined and cannot be overridden, although users can interact with
them directly through function arguments. Similarly, the image reading
and resizing module is fixed but will adapt to different Keras models by
checking their input shapes. The remaining seven modules are designed to
be overridable, enabling users to tailor the infrastructure to their
specific needs. These modules are discussed in detail on the software's
website \url{https://tengmcing.github.io/autovi/}.

@@

The \texttt{autovi.web} shiny application extends the functionality of
\texttt{autovi} by offering a user-friendly web interface for automated
residual plot assessment. This eliminates the common challenges
associated with software installation, so users can avoid managing
Python environments or handling version requirements for R libraries.
The platform is cross-platform and accessible on various devices and
operating systems, making it suitable even for users without R
programming experience. Additionally, updates are managed centrally,
ensuring that users always have access to the latest features. This
section discusses the implementation based on \texttt{autovi.web}
version 0.1.0.

@@

The interface \texttt{autovi.web} is built using the \texttt{shiny}
\citep{shiny} and \texttt{shinydashboard} \citep{shinydashboard} R
packages. Hosted on the \href{https://www.shinyapps.io}{shinyapps.io}
domain, the application is accessible through any modern web browser.
The R packages \texttt{htmltools} \citep{htmltools} and
\texttt{shinycssloaders} \citep{shinycssloaders} are used to render
markdown documentation in shiny application, and for loading animations
for shiny widgets, respectively.

@@

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-autovi-web-setup-1.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-autovi-web-lineup-1.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{autovi_paper_files/figure-pdf/fig-autovi-web-distributions-1.png}

@@

This paper presents new regression diagnostics software, the R package
\texttt{autovi} and its accompanying web interface, \texttt{autovi.web}.
It addresses a critical gap in the current landscape of statistical
software. While regression tools are widely available, effective and
efficient diagnostic methods have lagged behind, particularly in the
field of residual plot interpretation.

@@

The current version of \texttt{autovi} can be installed from CRAN, and
source code for both packages are available at
\url{https://github.com/TengMCing/autovi} and
\url{https://github.com/TengMCing/autovi_web} respectively. The web
interface is available from \url{autoviweb.netlify.app}.
